\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-17 paper/paper.Rnw
Updates

In this research, we had three basic factors:

@@

Additionally, there was a fourth factor for  = 48, 96:

@@

$$Power_{i} = \psi\left\{ \beta_{i,0} + \beta_{i,1} \cdot |\alpha| + \beta_{i,2}\cdot Sign + \beta_{i,3}\cdot Smoothed + \beta_{i,4}\cdot t + \beta_{i,5}\cdot |\alpha| \cdot Sign + \gamma_{i,1} \cdot subject_{j} \right\}$$

Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $\alpha$ is the true plot correlation, Sign is the direction of $\alpha$, $i$  is the plot design and $j$ is an observer specific random effect.

\begin{table}[htp]
\caption{Mixed Effects Logit Model for the predicted power of plot design. Sign = 1 if correlation > 0, 0 otherwise. Each plot design is estimated separately. Results must be transformed via the logit link function to generate a power prediction. Subjects with less than ten lineup evaluations were removed from the sample. Sample size for scatter plot model: 841, overlaid lines model: 843. The model is fitted using the lme4 package in R.}
\label{tab1}
\centering
<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
#full.sub$rabs <- factor(full.sub$rabs)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
full.sub$alpha <- log(.05/0.95)
#full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
#weirw.csv(full.sub, "stata.csv") #exporting data to other programs
m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
#pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
#pander(summary(m8s)$coefficients, justify=c('left', 'right', 'right', 'right', 'right'), caption="Model fit for scatter plots")
#library(xtable)
line.fit <- summary(m8l)$coefficients
scatter.fit <- summary(m8s)$coefficients
colnames(scatter.fit)[4] <- "p-value"
rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smoothed", "t", "Interaction(r+, r-)")
#colnames(line.fit)[4] <- "p-value"
#rownames(line.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
@
\end{table}

>>>>

The lineups were generated with three basic factors:

@@

Additionally, there was a fourth factor for $t = 48, 96$:

@@

$$Power_{ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma_{1} \cdot z_{1,s} \right\}$$

Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$ $x_{2,j} is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$

\begin{table}[htp]
\caption{Weighted Mixed Effects Logit Model for the predicted power of plot design. Sample size for scatter plot model: 900, overlaid lines model: 949, total: 1849. Default plot design is an overlaid lines type. The model is fitted using the glimmix procedure in SAS. Subjects with less than ten lineup evaluations were removed from the sample. %Sign = 1 if correlation > 0, 0 otherwise. Each plot design is estimated separately. Results must be transformed via the logit link function to generate a power prediction. Subjects with less than ten lineup evaluations were removed from the sample..}
\label{tab1}
\centering
<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
#full.sub$rabs <- factor(full.sub$rabs)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
full.sub$alpha <- log(.05/0.95)
#full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
#weirw.csv(full.sub, "choices.csv") #exporting data to be estimated in SAS.
#m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
#m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
#pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
#pander(summary(m8s)$coefficients, justify=c('left', 'right', 'right', 'right', 'right'), caption="Model fit for scatter plots")
#library(xtable)
#line.fit <- summary(m8l)$coefficients
#scatter.fit <- summary(m8s)$coefficients
#colnames(scatter.fit)[4] <- "p-value"
#rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smoothed", "t", "Interaction(r+, r-)")
#colnames(line.fit)[4] <- "p-value"
#rownames(line.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
#kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
#kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
@
\end{table}
